---
title: "YET_2605_Vis"
author: "FBender"
date: "2024-05-26"
output:
  word_document: default
---

# Loading libraries
```{r}
library(tidyverse)
library(ggplot2)
library(ggforce)
library(tidyr)
library(dplyr)
```

# Reading and merging Data
```{r}
# List all CSV files in the directory
csv_files <- list.files(path = "Data", pattern = "*.csv", full.names = TRUE)

yet_coltypes <- cols(
  Part = col_double(),
  Stim = col_character(),
  time = col_double(),
  x = col_double(),
  y = col_double(),
  x_pro = col_double(),
  y_pro = col_double(),
)

read_yet <- function(file) {
  read_csv(file, col_types = yet_coltypes)
}

# Read each CSV file into a data frame and store in a list
experiment_list <- 
  csv_files %>%
  map(~ read_yet(.x))

# Combine all data frames into one data frame
Experiment <- bind_rows(experiment_list)

# Read Stimuli and Conditions data
Stimuli <- 
  read_csv("AccuracyExperiment/Stimuli/Stimuli.csv") |> 
  rename(target_x = x, target_y = y)
  
  
Conditions <- read_csv("AccuracyExperiment/Conditions.csv")

# Merge Experiment data with Stimuli data
merged_data <- Experiment %>%
  left_join(Stimuli, by = c("Stim" = "File"))

# Merge the resulting data with Conditions data
final_data <- merged_data %>%
  left_join(Conditions, by = c("Part" = "Trial"))
```
Setting Screen parameters
```{r}
# Screen parameters
screen_diag_cm <- 40.64
screen_res_x <- 1920
screen_res_y <- 1080
viewing_distance_cm <- 45

# Calculate screen dimensions
aspect_ratio <- 16 / 9
screen_width_cm <- screen_diag_cm * (16 / sqrt(16^2 + 9^2))   
screen_height_cm <- screen_diag_cm * (9 / sqrt(16^2 + 9^2))

# Define the center of the screen
screen_center_x <- screen_res_x / 2
screen_center_y <- screen_res_y / 2
```


```{r}
# convert radians to degrees (atan)
rad2deg <- function(rad) {
  rad * (180 / pi)
}

# Function to calculate distance between two points in pixels
calculate_distance_px <- function(x1, y1, x2, y2) {
  sqrt((x2 - x1)^2 + (y2 - y1)^2)
}


# Function to calculate visual angle relative to the center of the screen
calculate_visual_angle_center <- function(x, y, center_x, center_y, screen_width_cm, screen_height_cm, screen_res_x, screen_res_y, viewing_distance_cm) {
  delta_x <- x - center_x
  delta_y <- y - center_y
  pixel_size_cm <- screen_width_cm / screen_res_x  
  
  angle_x <- rad2deg(2 * atan((delta_x * pixel_size_cm) / (2 * viewing_distance_cm)))       # calculates the horizontal visual angle (in degrees) of a gaze point relative to the center of the screen, based on its distance from the center, the pixel size, and the viewing distance
  angle_y <- rad2deg(2 * atan((delta_y * pixel_size_cm) / (2 * viewing_distance_cm)))       #  calculates the vertical visual angle (in degrees) of a gaze point relative to the center of the screen, based on its distance from the center, the pixel size, and the viewing distance
  
  return(list(angle_x, angle_y))
}

# Adding visual angle to the data
final_data <- final_data %>%
  mutate(
    angles = map2(x, y, ~calculate_visual_angle_center(.x, .y, screen_center_x, screen_center_y, screen_width_cm, screen_height_cm, screen_res_x, screen_res_y, viewing_distance_cm)),
    visual_angle_x = map_dbl(angles, 1),
    visual_angle_y = map_dbl(angles, 2)
  ) %>%
  select(-angles)  # Remove the intermediate column

# Function to calculate visual angle from distance in pixels
calculate_visual_angle_distance <- function(distance_px, screen_width_cm, screen_res_x, viewing_distance_cm) {
  pixel_size_cm <- screen_width_cm / screen_res_x  
  distance_cm <- distance_px * pixel_size_cm
  visual_angle <- rad2deg(2 * atan((distance_cm / 2) / viewing_distance_cm))
  return(visual_angle)
}

# Perform mutation on the entire data using rowwise()
final_data <- final_data %>%
  rowwise() %>%
  mutate(
    distance_px = calculate_distance_px(x, y, target_x, target_y),
    visual_angle_accuracy = calculate_visual_angle_distance(distance_px, screen_width_cm, screen_res_x, viewing_distance_cm)
  )

save(final_data, file = "FB24.rda")
```




# table with mean, median, and std for visual angle accuracy (in degrees)
```{r}
###### table with mean, median, and std for visual angle accuracy (in degrees)
##### C1 = IR, C2 = EC_L1, C3 = EC_L0
##### B1 = congruentBG(white), B2 = incongruentBG(black)

# Calculate required statistics including the background
stats <- final_data %>%
  group_by(Participant, Condition, background) %>%
  summarise(
    mean = mean(visual_angle_accuracy, na.rm = TRUE),
    median = median(visual_angle_accuracy, na.rm = TRUE),
    std = sd(visual_angle_accuracy, na.rm = TRUE),
    .groups = 'drop'
  )

# Reshape data to wide format with separate columns for each background
stats_wide <- stats %>%
  pivot_wider(
    names_from = c(Condition, background),
    values_from = c(mean, median, std),
    names_glue = "C{Condition}_B{background}_{.value}"
  )

# Reorganize columns
final_table <- stats_wide %>%
  select(
    Participant,
    C1_B1_mean, C1_B1_median, C1_B1_std,
    C1_B2_mean, C1_B2_median, C1_B2_std,
    C2_B1_mean, C2_B1_median, C2_B1_std,
    C2_B2_mean, C2_B2_median, C2_B2_std,
    C3_B1_mean, C3_B1_median, C3_B1_std,
    C3_B2_mean, C3_B2_median, C3_B2_std
  )
```


```{r}
##### same table, only for background 1 

# Filter for background 1
stats_background1 <- stats %>%
  filter(background == 1)

# Reorganize columns
final_table_b1 <- stats_wide %>%
  select(
    Participant,
    C1_B1_mean, C1_B1_median, C1_B1_std,
    C2_B1_mean, C2_B1_median, C2_B1_std,
    C3_B1_mean, C3_B1_median, C3_B1_std
  )
```

```{r}
##### table for background 1 excluding the outer stimuli on the left and right side (A1-4, F1-4)

# Stimuli to exclude
excluded_stimuli <- c(
  "24grid_WhiteA1.png", "24grid_WhiteA2.png", "24grid_WhiteA3.png", "24grid_WhiteA4.png",
  "24grid_WhiteF1.png", "24grid_WhiteF2.png", "24grid_WhiteF3.png", "24grid_WhiteF4.png"
)

# Calculate required statistics excluding excluded stimuli and including the background
stats <- final_data %>%
  filter(!Stim %in% excluded_stimuli) %>% # Exclude specific stimuli
  group_by(Participant, Condition, background) %>%
  summarise(
    mean = mean(visual_angle_accuracy, na.rm = TRUE),
    median = median(visual_angle_accuracy, na.rm = TRUE),
    std = sd(visual_angle_accuracy, na.rm = TRUE),
    .groups = 'drop'
  )

# Filter for background 1
stats_background1 <- stats %>%
  filter(background == 1)

# Reshape data to wide format with separate columns for each condition
stats_wide <- stats_background1 %>%
  pivot_wider(
    names_from = Condition,
    values_from = c(mean, median, std),
    names_glue = "C{Condition}_{.value}"
  )

# Reorganize columns
final_table_b1_exclAF <- stats_wide %>%
  select(
    Participant,
    C1_mean, C1_median, C1_std,
    C2_mean, C2_median, C2_std,
    C3_mean, C3_median, C3_std
  )
```

```{r}
#### table for background 1 now excluding all stimuli except for the 4 in the center

# Stimuli to exclude
excluded_stimuli <- c(
  "24grid_WhiteA1.png", "24grid_WhiteA2.png", "24grid_WhiteA3.png", "24grid_WhiteA4.png",
  "24grid_WhiteF1.png", "24grid_WhiteF2.png", "24grid_WhiteF3.png", "24grid_WhiteF4.png",
  "24grid_WhiteB1.png", "24grid_WhiteB2.png", "24grid_WhiteB3.png", "24grid_WhiteB4.png",
  "24grid_WhiteE1.png", "24grid_WhiteE2.png", "24grid_WhiteE3.png", "24grid_WhiteE4.png",
  "24grid_WhiteC1.png", "24grid_WhiteC4.png", 
  "24grid_WhiteD1.png", "24grid_WhiteD4.png"
)

# Calculate required statistics excluding excluded stimuli and including the background
stats <- final_data %>%
  filter(!Stim %in% excluded_stimuli) %>% # Exclude specific stimuli
  group_by(Participant, Condition, background) %>%
  summarise(
    mean = mean(visual_angle_accuracy, na.rm = TRUE),
    median = median(visual_angle_accuracy, na.rm = TRUE),
    std = sd(visual_angle_accuracy, na.rm = TRUE),
    .groups = 'drop'
  )

# Filter for background 1
stats_background1 <- stats %>%
  filter(background == 1)

# Reshape data to wide format with separate columns for each condition
stats_wide <- stats_background1 %>%
  pivot_wider(
    names_from = Condition,
    values_from = c(mean, median, std),
    names_glue = "C{Condition}_{.value}"
  )

# Reorganize columns
final_table_b1_center4 <- stats_wide %>%
  select(
    Participant,
    C1_mean, C1_median, C1_std,
    C2_mean, C2_median, C2_std,
    C3_mean, C3_median, C3_std
  )
```
# table with mean, median, and std for visual angle accuracy (in degrees)
## C1 = IR, C2 = EC_L1, C3 = EC_L0
## B1 = congruentBG(white), B2 = incongruentBG(black)
```{r}
print(final_table)
```
# same table, only for background 1
```{r}
print(final_table_b1)
```
# table for background 1 excluding the outer stimuli on the left and right side (A1-4, F1-4)
```{r}
print(final_table_b1_exclAF)
```
# table for background 1 now excluding all stimuli except for the 4 in the center
```{r}
print(final_table_b1_center4)
```




# Boxplots Mean and SD of Visual Angle Accuracy by Condition and Background 
```{r}
###### Meand and SD of Visual Angle Accuracy by Condition and Background 

# Group the data by condition and background
grouped_data <- final_data %>%
  group_by(Condition, background) %>%
  summarize(
    mean_visual_angle_accuracy = mean(visual_angle_accuracy, na.rm = TRUE),
    sd_visual_angle_accuracy = sd(visual_angle_accuracy, na.rm = TRUE),
    .groups = 'drop'
  )

# Print the summarized data
print(grouped_data)

ggplot(grouped_data, aes(x = factor(Condition), y = mean_visual_angle_accuracy, fill = factor(background))) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), width = 0.7) +
  geom_errorbar(aes(ymin = mean_visual_angle_accuracy - sd_visual_angle_accuracy, ymax = mean_visual_angle_accuracy + sd_visual_angle_accuracy), 
                position = position_dodge(width = 0.9), width = 0.25) +
  labs(title = "Mean and SD of Visual Angle Accuracy by Condition and Background",
       x = "Condition",
       y = "Mean Visual Angle Accuracy (in degrees)",
       fill = "Background") +
  theme_minimal()
```

# Mean and SD for Visual Angle Accuracy by Condition for congruent Background (1, white)
```{r}
###### Mean and SD for Visual Angle Accuracy by Condition for congruent Background (1, white)
###### Plots are arranged in the same order as the Stimuli shown on screen

# Group the data by condition, background, and Stim
grouped_data2 <- final_data %>%
  group_by(Condition, background, Stim) %>%
  summarize(
    mean_visual_angle_accuracy = mean(visual_angle_accuracy, na.rm = TRUE),
    sd_visual_angle_accuracy = sd(visual_angle_accuracy, na.rm = TRUE),
    .groups = 'drop'
  )

# Print the summarized data
print(grouped_data2)

# Filter the data to include only background 1
filtered_data <- grouped_data2 %>%
  filter(background == 1)

# Extract the number from the Stim names and create a new column
filtered_data <- filtered_data %>%
  mutate(Stim_number = as.numeric(str_extract(Stim, "\\d+(?=\\.png)")))

# Reorder the Stim variable based on the extracted number
filtered_data <- filtered_data %>%
  mutate(Stim = fct_reorder(Stim, Stim_number))

###### Mean and SD for Visual Angle Accuracy by Condition for congruent Background (1, white)
###### Plots are arranged in the same order as the Stimuli shown on screen
ggplot(filtered_data, aes(x = factor(Condition), y = mean_visual_angle_accuracy, fill = factor(background))) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), width = 0.7) +
  geom_errorbar(aes(ymin = mean_visual_angle_accuracy - sd_visual_angle_accuracy, ymax = mean_visual_angle_accuracy + sd_visual_angle_accuracy), 
                position = position_dodge(width = 0.9), width = 0.25) +
  labs(title = "Mean and SD of Visual Angle Accuracy by Condition for congruent Background (1. white)",
       x = "Condition",
       y = "Mean Visual Angle Accuracy") +
  theme_minimal() +
  guides(fill = FALSE) +
  facet_wrap(~ Stim, nrow = 4)

```

# inverted scatterplot showing stimulus positions and collected gaze points
```{r}
###### inverted scatterplot showing stimulus positions and collected gaze points for each background (inverted without affecting)
###### since 0,0 is in the upper left corner of the screen, I subtracted the screen height for the coordinates (without the stimuli and gaze points would be shown upside down)


create_scatterplot <- function(data, bg_value, screen_res_y) {
  plot_data <- data %>%
    mutate(y.x_plot = screen_res_y - y.x, 
           y.y_plot = screen_res_y - y.y)
  
  ggplot(data = plot_data, aes(x = x.y, y = y.y_plot)) +
    geom_point(data = filter(plot_data, background == bg_value & Condition == 1), aes(x = x.x, y = y.x_plot, color = "Condition 1"), size = 1) +
    geom_point(data = filter(plot_data, background == bg_value & Condition == 2), aes(x = x.x, y = y.x_plot, color = "Condition 2"), size = 1) +
    geom_point(data = filter(plot_data, background == bg_value & Condition == 3), aes(x = x.x, y = y.x_plot, color = "Condition 3"), size = 1) +
    geom_point(data = filter(plot_data, background == bg_value), aes(color = "Stimulus"), size = 3, alpha = 0.5) +
    scale_color_manual(name = "Points", values = c("Stimulus" = "black", "Condition 1" = "red", "Condition 2" = "blue", "Condition 3" = "green")) +
    labs(title = paste("Scatterplot for background =", bg_value),
         x = "X Coordinate", y = "Y Coordinate") +
    theme_minimal()
}
```
# inverted scatterplot showing stimulus positions and collected gaze points (bg1)
```{r}
# Scatterplot for background = 1
scatterplot_bg1 <- create_scatterplot(data = final_data, bg_value = 1, screen_res_y = screen_res_y)
print(scatterplot_bg1)
```
# inverted scatterplot showing stimulus positions and collected gaze points (bg2)
```{r}
# Scatterplot for background = 2
scatterplot_bg2 <- create_scatterplot(data = final_data, bg_value = 2, screen_res_y = screen_res_y)
print(scatterplot_bg2)
```


```{r}
############ Scatterplots for each participant

# Function to create scatterplot
create_scatterplot2 <- function(data, bg_value, screen_res_y) {
  plot_data <- data %>%
    mutate(y.x_plot = screen_res_y - y.x, 
           y.y_plot = screen_res_y - y.y)
  
  ggplot(data = plot_data, aes(x = x.y, y = y.y_plot)) +
    geom_point(data = filter(plot_data, background == bg_value & Condition == 1), aes(x = x.x, y = y.x_plot, color = "Condition 1"), size = 1) +
    geom_point(data = filter(plot_data, background == bg_value & Condition == 2), aes(x = x.x, y = y.x_plot, color = "Condition 2"), size = 1) +
    geom_point(data = filter(plot_data, background == bg_value & Condition == 3), aes(x = x.x, y = y.x_plot, color = "Condition 3"), size = 1) +
    geom_point(data = filter(plot_data, background == bg_value), aes(color = "Stimulus"), size = 3, alpha = 0.5) +
    scale_color_manual(name = "Points", values = c("Stimulus" = "black", "Condition 1" = "red", "Condition 2" = "blue", "Condition 3" = "green")) +
    labs(title = paste("Scatterplot for background =", bg_value),
         x = "X Coordinate", y = "Y Coordinate") +
    theme_minimal() +
    facet_wrap(~ Participant)  # Facet by Participant
}
```

```{r}
# Scatterplot for background = 1 with facets for each participant
scatterplot2_bg1 <- create_scatterplot2(data = final_data, bg_value = 1, screen_res_y = screen_res_y)
print(scatterplot2_bg1)
```

```{r}
# Scatterplot for background = 2 with facets for each participant
scatterplot2_bg2 <- create_scatterplot2(data = final_data, bg_value = 2, screen_res_y = screen_res_y)
print(scatterplot2_bg2)
```


# Stim positions with mean euclidian distance circles around them 
```{r}
####### Stim positions with mean euclidian distance circles around them 

# Function to calculate Euclidean distance between two points in pixels
calculate_euclidean_distance <- function(x1, y1, x2, y2) {
  distance <- sqrt((x2 - x1)^2 + (y2 - y1)^2)
  return(distance)
}

# Perform mutation on the entire data using rowwise()
final_data <- final_data %>%
  rowwise() %>%
  mutate(
    euclidean_distance = calculate_euclidean_distance(x.x, y.x, x.y, y.y)
  )

# Group the data by both Stim and Condition
final_data_G <- final_data %>%
  group_by(Stim, Condition) %>%
  # Calculate the mean Euclidean distance per stimulus per condition
  mutate(mean_euclidean_distance_per_stimulus_per_condition = mean(euclidean_distance)) %>%
  ungroup()  # Remove grouping

# Create a new dataframe for plotting with adjusted y-coordinates
plot_data <- final_data_G %>%
  mutate(
    y.x_plot = screen_res_y - y.x,
    y.y_plot = screen_res_y - y.y
  )

# Filter data for background 1 for plotting
plot_data_bg1 <- plot_data %>%
  filter(background == 1)

# Create the plot
ggplot(plot_data_bg1, aes(x = x.y, y = y.y_plot)) +
  geom_point(color = "black") +  # Black dots for stimulus positions
  geom_circle(data = plot_data_bg1 %>% filter(Condition == 1), aes(x0 = x.y, y0 = y.y_plot, r = mean_euclidean_distance_per_stimulus_per_condition, color = factor(Condition)), show.legend = TRUE, fill = NA, linewidth = 1.5) +  # Circles for condition 1
  geom_circle(data = plot_data_bg1 %>% filter(Condition == 2), aes(x0 = x.y, y0 = y.y_plot, r = mean_euclidean_distance_per_stimulus_per_condition, color = factor(Condition)), show.legend = TRUE, fill = NA, linewidth = 1.5) +  # Circles for condition 2
  geom_circle(data = plot_data_bg1 %>% filter(Condition == 3), aes(x0 = x.y, y0 = y.y_plot, r = mean_euclidean_distance_per_stimulus_per_condition, color = factor(Condition)), show.legend = TRUE, fill = NA, linewidth = 1.5) +  # Circles for condition 3
  scale_color_manual(name = "Condition", values = c("1" = "red", "2" = "blue", "3" = "green"), labels = c("Condition 1-IR", "Condition 2-EC_L1", "Condition 3-EC_L0")) +
  labs(x = "X Coordinate", y = "Y Coordinate", title = "Stimulus Positions with Mean Euclidean Distance Circles by Condition") +
  theme_minimal()
```

# Visual Angle Accuracy for horizontal and vertical shift
```{r}
########## Visual Angle Accuracy for horizontal and vertical shift

# Visual Angle Accuracy for horizontal shift (x) as points
ggplot(final_data, aes(x = visual_angle_x, y = visual_angle_accuracy, color = factor(background))) +
  geom_point() +
  labs(title = "Visual Angle Accuracy by Horizontal shift (x)",
       x = "Viewing Angle Horizontal (X)",
       y = "Visual Accuracy",
       color = "Background") +
  theme_minimal() +
  facet_grid(background ~ Condition, scales = "free_y")


# Visual Angle Accuracy for horizontal shift (x) as curves
ggplot(final_data, aes(x = visual_angle_x, y = visual_angle_accuracy, color = factor(background))) +
  geom_smooth(method = "loess", se = FALSE) +  # Fit a smooth curve
  labs(title = "Visual Angle Accuracy by Horizontal shift (x)",
       x = "Viewing Angle Horizontal (X)",
       y = "Visual Angle Accuracy",
       color = "Background") +
  theme_minimal() +
  facet_grid(background ~ Condition, scales = "free_y")


# Visual Angle Accuracy for vertical shift (y) as points
ggplot(final_data, aes(x = visual_angle_y, y = visual_angle_accuracy, color = factor(background))) +
  geom_point() +
  labs(title = "Visual Angle Accuracy by vertical shift (y)",
       x = "Viewing Angle Vertical (Y)",
       y = "Visual Angle Accuracy",
       color = "Background") +
  theme_minimal() +
  facet_grid(background ~ Condition, scales = "free_y")


# Visual Angle Accuracy for vertical shift (y) as curves
ggplot(final_data, aes(x = visual_angle_y, y = visual_angle_accuracy, color = factor(background))) +
  geom_smooth(method = "loess", se = FALSE) +  # Fit a smooth curve
  labs(title = "Visual Angle Accuracy by vertical shift (y)",
       x = "Viewing Angle Vertical (Y)",
       y = "Visual Angle Accuracy",
       color = "Background") +
  theme_minimal() +
  facet_grid(background ~ Condition, scales = "free_y")
```

```{r}

########## Combined curves by increase of X (solid) or Y (dashed)

# Plot for increase in x.x and y.x combined
ggplot(final_data, aes(color = factor(background))) +
  geom_smooth(aes(x = visual_angle_x, y = visual_angle_accuracy), method = "loess", se = FALSE) +
  geom_smooth(aes(x = visual_angle_y, y = visual_angle_accuracy), method = "loess", se = FALSE, linetype = "dashed") +
  labs(title = "Visual Angle Accuracy by Increase of X or Y",
       x = "Viewing Angle",
       y = "Visual Angle Accuracy",
       color = "Background") +
  theme_minimal() +
  facet_grid(background ~ Condition, scales = "free_y")

```

```{r}

######### filtering the first 10 rows for every Stim for each Trial (Part)
######### by excluding the first 10 observations for every stim in every trial, the rise at the center (0°) falls
######### this shows that the rise in the center is related to the initial recordings where the participant is still looking at the center where the recalibration point was located

# Filter out the first 10 rows for each Part and Stim
filtered_data <- final_data %>%
  group_by(Part, Stim) %>%
  slice(-(1:10)) %>%
  ungroup()

# Plot for increase in x.x and y.x combined
ggplot(filtered_data, aes(color = factor(background))) +
  geom_smooth(aes(x = visual_angle_x, y = visual_angle_accuracy), method = "loess", se = FALSE) +
  geom_smooth(aes(x = visual_angle_y, y = visual_angle_accuracy), method = "loess", se = FALSE, linetype = "dashed") +
  labs(title = "Visual Angle Accuracy by Increase of X or Y",
       x = "Viewing Angle",
       y = "Visual Angle Accuracy",
       color = "Background") +
  theme_minimal() +
  facet_grid(background ~ Condition, scales = "free_y")
```


# Model
# stan_glmer(visual_angle_accuracy ~ background * Condition + 
#    (background * Condition | Participant) + 
#    (Condition | Stim))
```{r}
# Fit the mixed-effects model with the specified interactions and random effects
model <- stan_glmer(
  visual_angle_accuracy ~ background * Condition + 
    (background * Condition | Participant) + 
    (Condition | Stim),
  data = final_data,
  family = gaussian(), 
  chains = 4,          
  cores = parallel::detectCores(), # Use all available cores
  iter = 100           
)

# Summarize the model
summary(model)

# Extract and plot the fixed effects
plot(model, pars = c("background", "Condition", "background:Condition"))

# Check the model diagnostics
posterior_interval(model, prob = 0.95)
```



# Model2
# same as before, for filtered data (- first 10 observations)
```{r}
# Fit the mixed-effects model with the specified interactions and random effects
model <- stan_glmer(
  visual_angle_accuracy ~ background * Condition + 
    (background * Condition | Participant) + 
    (Condition | Stim),
  data = filtered_data,
  family = gaussian(), 
  chains = 4,          
  cores = parallel::detectCores(), # Use all available cores
  iter = 100           
)

# Summarize the model
summary(model)

# Extract and plot the fixed effects
plot(model, pars = c("background", "Condition", "background:Condition"))

# Check the model diagnostics
posterior_interval(model, prob = 0.95)
```


